Abstract

### 2025-05-30 output/Response_to_action_editor_reviewers.Rmd

However, to our understanding, the more precise interpretation of the item parameters that results from our approach indeed provides a better insight into how the results are used.
A multidimensional representation of the items provides benchmarks for comparing different latent trait levels, as discussed by @reckase_difficulty_1985.
Therefore, the more accurate such a representation is, the better the latent trait interpretation will be.

Please see our answer on p. <!-- TODO: Complete (3?) --> of this letter for an example of how the interpretation of the item parameters may change when considering the covariance structure.
<!-- TODO: Complete -->

###### What implications does your approach have for standard psychometric procedures like equating.

Fitting a single-group model where the latent dimensions are correlated affects the computation of the multidimensional parameters we propose here.
Given the scale underidentification in IRT, it is relatively common to fix the latent variances to 1, so it is unlikely that non-standard distributions affect these applications.
In this sense, we have mentioned in our discussion that "in many applications both [the covariance- and correlation-based] versions will be coincident, fortunately" (p. 31, lines 2-3, or p. <!-- TODO: Complete -->, lines <!-- TODO: Complete --> in the new version).
However, in multi-group procedures, such as equating and linking, non-unitary variances (in some of the groups at least) may indeed affect the resulting multidimensional parameters.
Additionally, it is worth noting that the correlation matrix may differ among groups, which can also have a substantial effect on the multidimensional parameter values.
This is relevant because, even if the item parameters are equal across groups, the covariance structure may differ; thus, the multidimensional parameters will not be the same.
It is our understanding that both common item design and common person design procedures can be applied soundly under this approach.
Despite the latent and the test space being (generally) represented in a different geometric basis, note that if one of them is known, the other one is determined.
This linkage between the two spaces, along with the model invariance, implies that any transformation in one space must be accompanied by the corresponding transformation in the other.
These relationships allow for both common-item and common-person design equating procedures, derived according to Reckase [-@reckase_multidimensional_2009, see Chapter 8], and do not change in our approach.
For instance, if the common item-design procedure is used (where the item parameters in Form 2 of a test are transformed to resemble as closely as possible those in Form 1), both the item parameters and the covariance matrix will be affected.
Although it is out of the scope of our paper to explore in depth the implications for these psychometric procedures, we have added a new paragraph to our discussion briefly explaining this fact and acknowledging our limitation.
This new paragraph reads:

> Making a distinction between the two spaces can also have a significant impact
on our understanding of MIRT applications that involve several latent spaces,
such as multi-group equating and linking.
Although the usual transformations among coordinate systems still apply
under this consideration
[see Chapter 8 of @reckase_multidimensional_2009,
for an exhaustive discussion on the topic],
the different geometric bases on which each of the two spaces is represented
must be considered.
However, as they are mutually dependent,
any transformation in any of the two spaces will imply
the inverse transformation in the other one,
thus preserving the relationship between the two spaces.
Nevertheless, using non-orthogonal rotation matrices for aligning the items
(in a common-item design procedure) or persons
[in a common-person one, see Chapter 9 in @reckase_multidimensional_2009]
may have additional implications;
for example, besides a rescaling and an origin translation, a transformation
as defined, e.g., in Equation (8.31) of @reckase_multidimensional_2009,
may imply a rotation that can be non-orthogonal,
thus affecting the relative alignment among the test space axes,
and subsequently among the latent space ones.
Moreover, even if the transformation in one space is scale-invariant
(i.e., the transform matrix has unitary diagonal elements),
it may induce a rescaling in the other one,
given by the diagonal elements of the inverse transform matrix.
These implications will require future exploration of
the relationships between the latent and test spaces
as mathematically separate entities.

Conversely, the same multidimensional parameters may correspond to different parameter models if the latent covariance structure is different across groups.
Although it would be interesting to explore whether it is more realistic to assume equal M2PL parameters across groups (and thus possibly different multidimensional parameters) or vice-versa, we believe that this is an empirical question that lies outside the scope of our current manuscript.

### 2025-05-30 output/Response_to_action_editor_reviewers.Rmd

> Making a distinction between the two spaces can also have a significant impact
on our understanding of MIRT applications that involve several latent spaces,
such as multi-group equating and linking.
Although the usual transformations among coordinate systems still apply
under this consideration
[see Chapter 8 of @reckase_multidimensional_2009,
for an exhaustive discussion on the topic],
the different geometric bases on which each of the two spaces is represented
must be considered.
However, as they are mutually dependent,
any transformation in any of the two spaces will imply
the inverse transformation in the other one,
thus preserving the relationship between the two spaces.
Nevertheless, using non-orthogonal rotation matrices for aligning the items
(in a common-item design procedure) or persons
[in a common-person one, see Chapter 9 in @reckase_multidimensional_2009]
may have additional implications;
for example, besides a rescaling and an origin translation, a transformation
as defined, e.g., in Equation (8.31) of @reckase_multidimensional_2009,
may imply a rotation that can be non-orthogonal,
thus affecting the relative alignment among the test space axes,
and subsequently among the latent space ones.
Moreover, even if the transformation in one space is scale-invariant
(i.e., the transform matrix has unitary diagonal elements),
it may induce a rescaling in the other one,
given by the diagonal elements of the inverse transform matrix.
These implications will require future exploration of
the relationships between the latent and test spaces
as mathematically separate entities.

Regarding the multidimensional parameters, it is clear that different covariance matrices may have a substantial effect on the multidimensional parameter values.
This is relevant for multi-group procedures because, even if the item parameters are equal, if the covariance structure differs across groups, the multidimensional parameters will not be the same.
Conversely, the same multidimensional parameters may correspond to different parameter models if the latent covariance structure is different across groups.
Although it could be interesting to explore whether it is more realistic to assume equal M2PL parameters across groups (and thus possibly different multidimensional parameters) or vice-versa, we believe that this is an empirical question that also lies outside the scope of our current manuscript.

### 2025-06-02 output/Main_text.Rmd
Correct text size in body in example tables
Makes the table fill in better the size of the page
Reorder tables and figures
Makes a better use of the space in the manuscript
Apply "Header 1" style to document title
The document title is styled "normal", so it is changed to an inline "Header 1" instead

---
title: Abstract
output:
  officedown::rdocx_document:
    tables:
      style: APA_Like_style
      caption:
        style: Image Caption
        sep: ' '
    plots:
      caption:
        sep: ' '
    mapstyles:
      Body Text: ['First Paragraph', 'Abstract']
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
keywords: |
  Multidimensional 2-Parameter Logistic model,
  multidimensional item discrimination, multidimensional item difficulty,
  item vector representation, test space
bibliography: [../www/Multidimensional-parameters-MCLM.bib, ../www/packages.bib]
csl:          ../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

@@

@@

```{r empirical-table-Tezza2018-out}
#| cache:   false
#| tab.id:  empirical-table-Tezza2018-out
#| tab.cap: |
#|   Agnostic and Covariance-Based Multidimensional Item Parameters in
#|   Tezza et al. (2018)
item_params_output
```

>>>>

---
output:
  officedown::rdocx_document:
    tables:
      style: APA_Like_style
      caption:
        style: Image Caption
        sep: ' '
    plots:
      caption:
        sep: ' '
    mapstyles:
      Body Text: ['First Paragraph', 'Abstract']
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
keywords: |
  Multidimensional 2-Parameter Logistic model,
  multidimensional item discrimination, multidimensional item difficulty,
  item vector representation, test space
bibliography: [../www/Multidimensional-parameters-MCLM.bib, ../www/packages.bib]
csl:          ../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

@@

# Abstract

@@

```{r empirical-table-Tezza2018-out}
#| cache:   false
#| tab.id:  empirical-table-Tezza2018-out
#| tab.cap: |
#|   Agnostic and Covariance-Based Multidimensional Item Parameters in
#|   Tezza et al. (2018)
item_params_output |>
  fontsize(size = 11, part = "body")
```
